Acceleration of plague outbreaks in the second pandemic\\\emph{Electronic \supp}

### 2020-07-11 supp.Rnw
corrected foldval and calculation of fleas-per-rat requirement

As a vector-borne disease, the rate of bubonic plague spread is primarily affected by the flea-rat ratio.
In the traditional Ross-MacDonald model for vector-borne epidemics, $\rzero$ is proportional to the square of vector-host ratio because vectors must independently bite hosts twice (once to become infected and a second time to infect a susceptible host). In the rat-flea model, $\rzero$ is instead proportional to $V/H$ because \emph{all} infected fleas on a host disperse and bite other hosts when their initial host dies.
From Keeling and Gilligan's plague model \cite{KeelGill00b}, the
expected change in $r$ for a change in flea-rat ratio from $K_{\rm
  F0}$ to $K_{\rm F}$ is $(K_{\rm F}/K_{\rm F0})^S$, where $S$ is the
sensitivity ($\approx 1.5$). Thus in order to a see a fivefold
\david{why FIVEfold?  we need only FOURfold; didn't we have a macro
  for this?} increase in $r$ we would need a change of $5^{(1/S)} \approx \Sexpr{round(5^(1/1.5),1)}$, not in the rat density, but in the number of fleas per rat, which seems unlikely.

>>>>

As a vector-borne disease, the rate of bubonic plague spread is primarily affected by the flea-rat ratio.
In the traditional Ross-MacDonald model for vector-borne epidemics, $\rzero$ is proportional to the square of vector-host ratio because vectors must independently bite hosts twice (once to become infected and a second time to infect a susceptible host). In the rat-flea model, $\rzero$ is instead proportional to $V/H$ because \emph{all} infected fleas on a host disperse and bite other hosts when their initial host dies.
From Keeling and Gilligan's plague model \cite{KeelGill00b}, the
expected change in $r$ for a change in flea-rat ratio from $K_{\rm
  F0}$ to $K_{\rm F}$ is $(K_{\rm F}/K_{\rm F0})^S$, where $S$ is the
sensitivity ($\approx 1.5$). Thus in order to a see a \foldval
increase in $r$ we would need a change of $4^{(1/S)} \approx \Sexpr{round(4^(1/1.5),1)}$, not in the rat density, but in the number of fleas per rat, which seems unlikely.
